---
title: "Spatial Data Visualization"
description: |
  A short description of the post.
author:
  - name: Marie
    url: {}
date: 03-01-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---

## California Oil Spill Spatial Data Visualization

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# Attach packages
library(tidyverse)
library(here)
library(janitor)
library(sf)
library(tmap)

# Read in CA Counties data
ca_counties <- read_sf(here("_posts", "2021-03-14-spatial-analysis", "ca_counties"), layer = "CA_Counties_TIGER2016") %>% 
  clean_names() %>% 
  select("name")

# Read in oil spill data
oil_spills <- read_sf(here("_posts", "2021-03-14-spatial-analysis", "oil_spill_shapefile"), layer = "Oil_Spill_Incident_Tracking") %>% 
  clean_names() %>% 
  relocate("dateofinci")

```

### **Figure 1.** Interactive oil spill map in tmap. 
```{r}

#---------------------------------------
# Make interactive map with tmap
#---------------------------------------

tmap_mode("view")

tm_basemap("Stamen.TerrainBackground") +
tm_shape(oil_spills) +
  tm_dots(col = "red", legend.show = FALSE) +
  tm_layout(title = "Oil spill locations and dates in California")

```
California oil spill locations shown in red, hover to show date.

***

### **Figure 2.** Chloreopleth Map of oil spills by county in ggplot.

```{r}

#---------------------------------------
# Make choropleth map with ggplot
#---------------------------------------

# Spatial join and calculate number of spills per county
ca_spills <- st_join(ca_counties, oil_spills) %>% 
  group_by(name) %>%
  summarize(
    counts = n()
  )

# Plot
ggplot() +
  geom_sf(data = ca_spills, aes(fill = counts)) +
  labs(fill = "Number of oil spills") +
  scale_fill_gradient(low = "#F9C48B", high = "#F25C54") +
  labs(title = "Number of oil spills per county in California") +
  theme_void()

```
<br>
Counties with the highest number of oil spill incidents are shown in \n orange and counties with the lowest overall incidents are shown in pale yellow.

***

**Data source:** Title Oil Spill Incident Tracking [ds394]
Publication date 2009-07-23. Edition 2008. Presentation formats digital map.
FGDC geospatial presentation format vector digital data. Available at: