---
title: "Text Analysis"
description: |
  Text analysis of Good Omens by Neil Gaiman and Terry Pratchett.
author:
  - name: Marie
    url: {}
date: 02-25-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach Packages
library(tidyverse)
library(here)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
library(png)
library(ggimage)
library(patchwork)

```
<center>
# Armageddon Only Happens Once: Good Omens Text Analysis

<br>

**Synopsis:** The world is going to end next Saturday, just before dinner, but it turns out there are a few problems — the Antichrist has been misplaced, the Four Horseman of the Apocalypse ride motorcycles and the representatives from heaven and hell decide that they like the human race. (Synopsis from [NPR](https://www.npr.org/books/titles/392587202/good-omens-the-nice-and-accurate-prophecies-of-agnes-nutter-witch))

</center>

<br>

## Word Counts Visualizations

**Figure 1.** Word Count Plot

<br>

```{r, cache = TRUE}

#---------------------------------------
# Text Wrangling
#---------------------------------------

# Read in Good Omens pdf
omens_text <- pdf_text(here("_posts/2021-02-25-text-analysis","Media","GoodOmens.pdf"))

# Tidy
omens_tidy <- data.frame(omens_text) %>% 
  mutate(text_full = str_split(omens_text, pattern = "\\n")) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) %>% 
  slice(-(63:109)) %>% # Skip character descriptions
 mutate(chapter = case_when(
    str_detect(text_full, "^In the beginning$|^Eleven years ago$|^Wednesday$|^Thursday$|^Friday$|^Saturday$|^Sunday$") ~ text_full,
    TRUE ~ NA_character_
  )) %>% 
  fill(chapter) %>% 
  mutate(chapter = fct_relevel(chapter, 
            "In the beginning", "Eleven years ago", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) %>%  # Set chapter order
  unnest_tokens(word, text_full) %>% # Tokenize
  mutate(word = str_replace(word, "[0-9-]+", NA_character_)) %>% # remove rogue page numbers
  select(-omens_text) %>% 
  drop_na() %>% 
  anti_join(stop_words) # Remove stop-words

# Get word counts of non stop-words
nonstop_counts <- omens_tidy %>% 
  count(chapter, word)

#---------------------------------------
# Visualize top word counts by chapter
#---------------------------------------

# Find top 5 words by chapter
top_5_words <- nonstop_counts %>% 
  group_by(chapter) %>% 
  arrange(-n) %>% 
  slice(1:5)

# Visualize top 5 words per chapter

omens_palette <- c("#A6D3DD", "#FDF3B0", "#F0B140", "#B4B4B4", "#C6333B", "#671D21", "#070707")

ggplot(data = top_5_words, aes(x = word, y = n)) +
  geom_col(aes(fill = chapter), show.legend = FALSE) +
  scale_fill_manual(values = omens_palette) +
  facet_wrap(~chapter, scales = "free") +
  theme_minimal() +
  coord_flip() +
  labs(title = "Five most common words per chapter in Good Omens",
       x = "Word",
       y = "Number of occurences") +
    theme(plot.title = element_text(color="#C6333B", 
                                  size=16,
                                  face = "bold",
                                  hjust = 0.5))

```

**Figure 2.** Word Cloud Plot

<br>

```{r}

#---------------------------------------
# Visualize top 100 words in book
#---------------------------------------

omens_top100 <- nonstop_counts %>% 
  arrange(-n) %>% 
  slice(1:100) %>% 
  mutate(angle = 90 * sample(c(0, 1), n(), replace = TRUE, prob = c(60, 40)))

omens_cloud <- ggplot(data = omens_top100, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, 
                          size = n,
                          angle = angle),
                          shape = "squares") +
  scale_size_area(max_size = 10) +
  scale_color_gradientn(colors = c(low = "#A6D3DD", middle = "#F0B140", high = "#C6333B")) +
  labs(title = "Top 100 Words in Good Omens") +
  theme(plot.title = element_text(color="#C6333B", 
                                  size = 16,
                                  face = "bold",
                                  hjust = 0.5))

# Add background
img = here("_posts/2021-02-25-text-analysis","Media","angel-demon-buds.png")

ggbackground(omens_cloud, img)


```

## Sentiment Analysis

**Figure 3.** Positive vs. Negative Sentiment Analysis

<br>

```{r}

#---------------------------------------
# Sentiment analysis with afinn
#---------------------------------------

# Join afinn lexicon
omens_afinn <- omens_tidy %>% 
  inner_join(get_sentiments("afinn"))

# Calculate mean score per chapter
afinn_means <- omens_afinn %>% 
  group_by(chapter) %>% 
  summarize(mean_afinn = mean(value))

# Visualize positive vs. negative sentiment per chapter
ggplot(data = afinn_means, 
       aes(x = fct_rev(as.factor(chapter)), 
           y = mean_afinn)) +
  geom_col(aes(fill = chapter), show.legend = FALSE) +
  scale_fill_manual(values = omens_palette) +
  theme_minimal() +
  labs(
    title = "Positive and negative sentiments by chapter\n of of Good Omens",
    x = "Chapter",
    y = "Average Sentiment"
  ) +
  theme(plot.title = element_text(color="#C6333B", 
                                  size = 16,
                                  face = "bold",
                                  hjust = 0.5)) +
  coord_flip()

```

**Figure 4.** Sentiment Bins Plot

<br>

```{r}

#---------------------------------------
# Sentiment analysis with nrc
#---------------------------------------

# join nrc lexicon
omens_nrc <- omens_tidy %>% 
  inner_join(get_sentiments("nrc"))

# Get counts
omens_nrc_counts <- omens_nrc %>% 
  count(chapter, sentiment)

# Visualize sentiment bins by chapter
ggplot(data = omens_nrc_counts, aes(x = sentiment, y = n)) +
  geom_col(aes(fill = chapter), show.legend = FALSE) +
  scale_fill_manual(values = omens_palette) +
  theme_minimal() +
  labs(
    title = "Top sentiment bins by chapter of Good Omens",
    x = "Chapter",
    y = "Number of occurences"
  ) +
  theme(plot.title = element_text(color="#C6333B", 
                                  size = 16,
                                  face = "bold",
                                  hjust = 0.5)) +
  facet_wrap(~chapter) +
  coord_flip()

```


***

**References:** <br>
Gaiman, N., & Pratchett, T. (1990). Good Omens: The Nice and Accurate Prophecies of Agnes Nutter, Witch. London: Berkley Books.

Mohammad, S. & Turney, P. (2013). Crowdsourcing a Word-Emotion Association Lexicon. Computational Intelligence, 29 (3), 436-465.

Nielsen, F.A. (2011). A new ANEW: Evaluation of a word list for sentiment analysis in microblogs. Proceedings of the ESWC2011 Workshop on 'Making Sense of Microposts': Big things come in small packages 718 in CEUR Workshop Proceedings 93-98.